New lower bounds for Schur and weak Schur numbers

### 2021-04-09 article/schur.tex

\begin{theorem}
	Let \((n,k), (p,q) \in (\mathbb{N}^*)^2\). If there exists a \(SF\)-template of \(k+1\) colors and lenght \(p\),
and a partition of \(n\) sum-free subsets of \([\![1,q]\!]\) then there exists a partition of \(n+k\) sum-free subsets
of \([\![1,pq]\!]\).
\end{theorem}

@@

\begin{corollary}
	Let \(n, k \in \mathbb{N}^*\), we have \\
	\[ S(n+k) \geqslant S^+(k+1)S(n) \]
\end{corollary}

See below an example of such a construction. \\
It is quite similar to Abott and Hanson construction \cite{AbbottHanson}. They extend vertically a sum-free partition,
and horizontally an other sum-free partition. This way each "block" acts like a security zone for the other one.
Here, the horizontal partition is no longer to the side of the vertical one, but occupies the column of the special color
of the \(SF\)-template, i.e the one without the extra condition.

@@

We now define \(h : [\![1,pq]\!] \longrightarrow [\![1,n+k]\!] \) as follows :
\( \forall x \in [\![1,pq]\!] \), we write \(x = \alpha p - u\) where \(\alpha \in [\![1,q]\!] \) and \( u \in
[\![0,p-1]\!] \).
This decomposition if of course unique. If \(f(p-u) \leqslant k\), we set \(h(x) = f(p-u)\), else (i.e \(f(p-u) = k + 1)\) \(h(x) = k + g(\alpha) \).
Prooving that the partition of \([\![1,pq]\!]\) induced by \(h\) is sum-free will complete the proof.

>>>>

\begin{theorem}
	Let \((n,k), (p,q) \in (\mathbb{N}^*)^2\). If there exists a \(SF\)-template of \(k+1\) colors and lenght \(p\),
and a partition of \(n\) sum-free subsets of \([\![1,q]\!]\) then there exists a partition of \(n+k\) sum-free subsets
of \([\![1,pq+m_{k+1}-1]\!]\). \(m_{k+1}\) is the first number colored with the \(k+1\)-th color in the \(SF\)-template.
\end{theorem}

@@

\begin{corollary}
	Let \(n, k \in \mathbb{N}^*\), we have \\
	\[ S(n+k) \geqslant S^+(k+1)S(n) + m_{k+1} - 1 \]
\end{corollary}

The idea lying beneath this inequality is similar to Abbott and Hanson's contruction \cite{AbbottHanson}. 
They extend vertically a sum-free partition, and horizontally an other sum-free partition. This way each "block" 
acts like a security zone for the other one. Here, the horizontal partition is no longer to the side of the 
vertical one, but occupies the column of the special color of the \(SF\)-template, i.e the one without the 
extra condition. We give the following table for \(p = 7, q = 4, n = 2 ~\textrm{and}~ k = 2\) to make the intuition clear. \\

This shows the inequality \(S(2+2) \geqslant S^+(3)S(2)\), with the extra constant. 
We prooved \(S^+(3) = 9\).The special color is blue.

\(\overset{SF\textrm{-template}}{\textrm{Sum-free partition} \left \updownarrow\overleftrightarrow{\begin{tabular}{|*{9}{c|}}
	\hline
	\cellcolor{red} 1 & \cellcolor{green} 2 & \cellcolor{green} 3 & \cellcolor{red} 4 & \cellcolor{blue} 5 & \cellcolor{blue} 6 & \cellcolor{red} 7 & \cellcolor{blue} 8 & \cellcolor{blue} 9\\
	\hline
	\cellcolor{red} 10 & \cellcolor{green} 11 & \cellcolor{green} 12 & \cellcolor{red} 13 & \cellcolor{yellow} 14 & \cellcolor{yellow} 15 & \cellcolor{red} 16 & \cellcolor{yellow} 17 & \cellcolor{yellow} 18\\
	\hline
	\cellcolor{red} 19 & \cellcolor{green} 20 & \cellcolor{green} 21 & \cellcolor{red} 22 & \cellcolor{yellow} 23 & \cellcolor{yellow} 24 & \cellcolor{red} 25 & \cellcolor{yellow} 26 & \cellcolor{yellow} 27\\
	\hline
	\cellcolor{red} 28 & \cellcolor{green} 29 & \cellcolor{green} 30 & \cellcolor{red} 31 & \cellcolor{blue} 32 & \cellcolor{blue} 33 & \cellcolor{red} 34 & \cellcolor{blue} 35 & \cellcolor{blue} 36\\
	\hline
	\cellcolor{red} 37 & \cellcolor{green} 38 & \cellcolor{green} 39 & \cellcolor{red} 40 & & & & &\\
	\hline
\end{tabular}}
\right.}\)
 \\

\begin{center}
\(SF\)-template with 3 colors (it is actually maximal)\\
\begin{tabular}{|*{9}{c|}}
	\hline 
	\cellcolor{red} 1 & \cellcolor{green} 2 & \cellcolor{green} 3 & \cellcolor{red} 4 & \cellcolor{blue} 5 & \cellcolor{blue} 6 & \cellcolor{red} 7 & \cellcolor{blue} 8 & \cellcolor{blue} 9\\
	\hline
\end{tabular} \\

Sum-free partition (\(S(2)\) actually) \\
\begin{tabular}{|*{4}{c|}}
	\hline 
	\cellcolor{blue} 1 & \cellcolor{yellow} 2 & \cellcolor{yellow} 3 & \cellcolor{blue} 4 \\
	\hline
\end{tabular}
\end{center}

To compare, here is a classic Abbott and Hanson partition : \(S(2+2) \geqslant S(2)(2S(2)+1)+S(2) \). \\

\begin{center}
\begin{tabular}{|*{9}{c|}}
	\hline
	\cellcolor{red} 1 & \cellcolor{green} 2 & \cellcolor{green} 3 & \cellcolor{red} 4 & \cellcolor{blue} 5 & \cellcolor{blue} 6 & \cellcolor{blue} 7 & \cellcolor{blue} 8 & \cellcolor{blue} 9\\
	\hline
	\cellcolor{red} 10 & \cellcolor{green} 11 & \cellcolor{green} 12 & \cellcolor{red} 13 & \cellcolor{yellow} 14 & \cellcolor{yellow} 15 & \cellcolor{yellow} 16 & \cellcolor{yellow} 17 & \cellcolor{yellow} 18\\
	\hline
	\cellcolor{red} 19 & \cellcolor{green} 20 & \cellcolor{green} 21 & \cellcolor{red} 22 & \cellcolor{yellow} 23 & \cellcolor{yellow} 24 & \cellcolor{yellow} 25 & \cellcolor{yellow} 26 & \cellcolor{yellow} 27\\
	\hline
	\cellcolor{red} 28 & \cellcolor{green} 29 & \cellcolor{green} 30 & \cellcolor{red} 31 & \cellcolor{blue} 32 & \cellcolor{blue} 33 & \cellcolor{blue} 34 & \cellcolor{blue} 35 & \cellcolor{blue} 36 \\
	\hline
	\cellcolor{red} 37 & \cellcolor{green} 38 & \cellcolor{green} 39 & \cellcolor{red} 40 & & & & &\\
	\hline
\end{tabular}
\end{center}

@@

We now define \(h : [\![1,pq]\!] \longrightarrow [\![1,n+k]\!] \) as follows :
\( \forall x \in [\![1,pq + m_{k+1}-1]\!] \), we write \(x = \alpha p - u\) where \(\alpha \in [\![1,q]\!] \) and \( u \in
[\![0,p-1]\!] \).
This decomposition if of course unique. If \(f(p-u) \leqslant k\), we set \(h(x) = f(p-u)\), else (i.e \(f(p-u) = k + 1)\) \(h(x) = k + g(\alpha) \).
Prooving that the partition of \([\![1,pq]\!]\) induced by \(h\) is sum-free will complete the proof.
